---
title: "Lab 6 Report"
author: "Christopher Byrd"
date: "11/10/2024"
output:
  pdf_document: default
---

```{r include=FALSE, warning=FALSE}
library(plyr)
library(DescTools)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(hrbrthemes)
library(tidyverse)
library(viridis)
library(forcats)
library(readxl)
library(knitr)

Module_6_Lab_Data_1_ <- read_excel(
  "C:/Users/byrds/Downloads/Module 6 Lab Data (1).xlsx", 
  col_types = c("numeric"))

incomeData <- Module_6_Lab_Data_1_[[1]]
```

### Section 1: Hypothesis

The goal of this research paper is to determine whether the average 
income of people who pray regularly in the US is similar to that if the 
rest of the world. To that end, the null and research hypothesis are as 
follows:

> Null: There is no significant difference in income between regular 
prayers in the US and other countries.\
> Research: Those who are praying regularly in the US will have a higher 
average income than individuals in other countries.

> H~0~: USA~Income~ = World~Income~\
> H~1~: USA~Income~ > World~Income~

### Section 2: Data

The data used for this report is regarding the USA is from the 2017 
*Baylor Survey of Religious Life*. Participants selected an ordinal 
range describing their yearly income, with midpoints of each category 
compiled for research purposes. Data for other nations was ascertained 
from the internet.

### Section 3: Method

Using RStudio, the data set was packaged into a variable,
*`incomeData`*. From there, the following code was ran:

```{r results='hide'}
test01 <- ZTest(incomeData, mu = c(11230), sd_pop = c(679),
                alternative = c("greater"), paired = FALSE)
```
A right-handed test was performed to determine if the US has a higher 
average income, and, if so, whether it was statistically significant.\

```{r results='hide'}
format_ztest <- function(test) {
  sprintf("z = %.2f, StDev Pop = %.2f, p-value = %.4f*, 95%% CI = (%.2f, %.2f)",
          test$statistic, test$parameter, test$p.value,
          test$conf.int[1], test$conf.int[2])
}
```

### Section 4: Results

The results of the test is as follows:

```{r echo=FALSE, results='markup'}
cat(format_ztest(test01))
```

### Section 5: Discussion and Conclusion

From these results, the idea that those who are praying regularly in 
2017 are older than those in 2007 is statistically significant.
Because this, we can reject the null hypothesis and accept the research
hypothesis.\

What didn't fair so well was my previous analyses of the prayer habits 
of ages 21 and 69. According to the results, there was no significant 
difference in observed ages between samples in each age group.
Accordingly, in relation to this group, we can accept the null, even
if the research hypothesis isn't quite applicable, as there is no
significant changes within the average age in each group. This means
that any increase or decrease in chance may be caused by sampling
error or chance.