---
title: "Unit 9 Lab Report"
author: "Christopher Byrd"
date: '`r Sys.Date()`'
output:
  pdf_document:
    latex_engine: lualatex
    highlight: pygments
    toc: true
    toc_depth: 3
mainfont: Times New Roman
---

```{r setupChunk, warning=FALSE, include=FALSE}
library(plyr)
library(DescTools)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(hrbrthemes)
library(tidyverse)
library(viridis)
library(forcats)
library(readxl)
library(knitr)
library(addinslist)
library(clipr)
library(gmodels)
library(thematic)

Module_9_Lab_Data <- read_excel("data/Module 9 Lab Data.xlsx", 
    col_types = c("numeric", "numeric"))
```

### Section 1: Hypothesis

The goal of this research paper is to determine if regular prayer has
any correlation on the opinion of the federal government allowing prayer 
in school. The following are the hypothesizes examined:

> Null: There is no significant correlation between prayer frequency and
> opinion on prayer in school.\
> Research~1~: There will be a significance between prayer frequency and
> opinion on prayer in schools.

> H~0~: Higher prayer frequency ≠ Significance on opinion of prayer in
School\
> H~1~: Higher prayer frequency = Significance on opinion of prayer in 
School

### Section 2: Data

The data used for this report is from the Baylor Survey of Religious
Life. The answers of participants are encoded as follows:

\newpage
| Question asked or description      |         Answers          |
|------------------------------------|--------------------------|
| About how much time do you         | 0 - Never                |
| spend praying alone outside        | 1 - On certain occasions |
| of religious services?             | 2 - Once a week or less  |
|                                    | 3 - A few times a week   |
|                                    | 4 - Once daily           |
|                                    | 5 - Several times a day  |
|------------------------------------|--------------------------|
| The federal government should      | 1 - Strongly disagree    |
| allow prayer in public school      | 2 - Disagree             |
|                                    | 3 - Agree                |
|                                    | 4 - Strongly agree       |
|                                    | 8 - Undecided            |

The preceding data is nominal.

### Section 3: Method

Using RStudio, the data was imported from an excel sheet using the
`readxl` package and directed into the variable `Module_9_Lab_Data`. To
analyse the data, the function `CrossTable()` was run. It creates a
cross tabulation table and runs a chi-squared test.

```{r warning=FALSE, results='hide', eval=FALSE}
CrossTable(Module_9_Lab_Data$PrayerFreq, Module_9_Lab_Data$PrayerSchool,
           expected = TRUE, dnn = c("Prayer Frequency",
                                    "Prayer in School"),
           format = c("SPSS"))
```

From here, a bar chart of opinions on prayer in school related to prayer
habits was created. To do so was a bit of a task. 

```{r warning=FALSE, results='hide', eval=FALSE}
data <- Module_9_Lab_Data

result <- data %>%
  group_by(PrayerFreq, PrayerSchool) %>%
  summarise(Count = n(), .groups = 'drop')

# View the result
print(result, n = 35)
```

To create the graph, prayer habit related to total occurrence of prayer
in school opinion has to be calculated. There isn't a completely 
straightforward way of doing this in R, so I opted to use summarise and
print to find the totals, as seen above. Through this, I learned `%>%` is
called the pipe symbol and feeds output to the subsequent function 
(essentially doing what `<-` does backwards).\

Then came the tedious part. I transcribed the amount of occurrences by
hand to a `data.frame` capable of displaying the data in the graph. I'm 
not entirely certain why I had to use `rep()` on my Habit variable, other
than RStudio threw an error saying that lengths between response and
habit must be identical.

```{r warning=FALSE, results='hide'}
# Data
data <- data.frame(
  Habit = rep(paste("Habit", 0:5), each = 5),
  Response = rep(c("Strongly Disagree", "Disagree",
                   "Agree", "Strongly Agree", "Undecided"), times = 6),
  Count = c(117, 46, 39, 7, 15,   # Habit 0
            41, 56, 98, 35, 26,   # Habit 1
            9, 13, 40, 13, 9,     # Habit 2
            14, 26, 83, 55, 14,   # Habit 3
            17, 25, 108, 97, 20,  # Habit 4
            17, 26, 112, 198, 19) # Habit 5  
)
```

Finally, here's the graph code, with one quick add-on: originally, the
responses were displayed without a logical order. They had to be manually
coerced to display in the order desired using a `factor()`.

```{r warning=FALSE, results='hide', eval=FALSE}
# Reordering Response levels
data$Response <- factor(data$Response, 
                        levels = c("Undecided", "Strongly Disagree",
                                   "Disagree", "Agree", 
                                   "Strongly Agree"))

# Create the plot
ggplot(data, aes(x = Response, y = Count, fill = Response)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ Habit, scales = "free_y") +
  labs(
    title = "Survey Responses for Prayer Habits",
    x = "Prayer should be allowed in public schools",
    y = "Count") +
  theme(
    axis.text.x = element_blank()
    ) +
  scale_fill_manual(values = c("#787878", "#ff1100",
                               "#ff6961", "#92ff8c", "#0dff00"))
```

Note that the y-axis are *intentionally* different in each graph in 
figure 1. This is to better illustrate relationships in the graphs
rather than just raw numbers. Figure 2 offers an un-adjusted view for
better transparency on raw numbers.

### Section 4: Results

The results from `CrossTable()` can be viewed in full in figure 3. As for
the results of the Chi-squared test (fig. 4), it returns a p-value of 
1.06E-89, well into statistical significance. This is further
demonstrated in figure 1.

### Section 5: Discussion and Conclusion

From these results, it's clear that prayer frequency has a direct and 
measurable impact on opinion regarding the allowance of prayer by the
federal government in schools.

It stands to reason the null is rejected and H~1~ is accepted.

### Section 6: Graphics and Visuals
*Note formatting in this section is weird due to figure sizes.*\

```{r echo=FALSE, fig.cap="Bar graph with adjusted Y-value. Opinion on prayer in public school based upon prayer habit."}
# Reordering Response levels
data$Response <- factor(data$Response, 
                        levels = c("Undecided", "Strongly Disagree", "Disagree", "Agree", "Strongly Agree"))

# Create the plot
ggplot(data, aes(x = Response, y = Count, fill = Response)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ Habit, scales = "free_y") +
  labs(
    title = "Survey Responses for Prayer Habits",
    x = "Prayer should be allowed in public schools",
    y = "Count") +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )+
  scale_fill_manual(values = c("#787878", "#ff1100",
                               "#ff7878", "#92ff8c", "#0dff00"))
```

```{r echo=FALSE, fig.cap="Bar graph with non-adjusted Y-value. Opinion on prayer in public school based upon prayer habit."}
# Reordering Response levels
data$Response <- factor(data$Response, 
                        levels = c("Undecided", "Strongly Disagree", "Disagree", "Agree", "Strongly Agree"))

# Create the plot
ggplot(data, aes(x = Response, y = Count, fill = Response)) +
  geom_bar(stat = "identity") +
  facet_wrap(~ Habit) +
  labs(
    title = "Survey Responses for Prayer Habits",
    x = "Prayer should be allowed in public schools",
    y = "Count") +
  theme_minimal() +
  theme(
    axis.text.x = element_blank(),
    axis.ticks.x = element_blank()
  )+
  scale_fill_manual(values = c("#787878", "#ff1100",
                               "#ff7878", "#92ff8c", "#0dff00"))
```

\newpage

|Prayer Opinion:  |        1  |        2  |        3  |        4  |        8  |           | 
|-----------------|----------:|----------:|----------:|----------:|----------:|----------:|
|Prayer Frequency |           |           |           |           |           | Row Total |
|-----------------|           |           |           |           |           |-----------|
|               0 |      117  |       46  |       39  |        7  |       15  |      224  | 
|                 |   34.523  |   30.830  |   77.075  |   65.032  |   16.539  |           | 
|                 |  197.038  |    7.464  |   18.809  |   51.786  |    0.143  |           | 
|                 |   52.232% |   20.536% |   17.411% |    3.125% |    6.696% |   16.057% | 
|                 |   54.419% |   23.958% |    8.125% |    1.728% |   14.563% |           | 
|                 |    8.387% |    3.297% |    2.796% |    0.502% |    1.075% |           | 
|-----------------|-----------|-----------|-----------|-----------|-----------|-----------|
|               1 |       41  |       56  |       98  |       35  |       26  |      256  | 
|                 |   39.455  |   35.234  |   88.086  |   74.323  |   18.902  |           | 
|                 |    0.060  |   12.238  |    1.116  |   20.805  |    2.666  |           | 
|                 |   16.016% |   21.875% |   38.281% |   13.672% |   10.156% |   18.351% | 
|                 |   19.070% |   29.167% |   20.417% |    8.642% |   25.243% |           | 
|                 |    2.939% |    4.014% |    7.025% |    2.509% |    1.864% |           | 
|-----------------|-----------|-----------|-----------|-----------|-----------|-----------|
|               2 |        9  |       13  |       40  |       13  |        9  |       84  | 
|                 |   12.946  |   11.561  |   28.903  |   24.387  |    6.202  |           | 
|                 |    1.203  |    0.179  |    4.260  |    5.317  |    1.262  |           | 
|                 |   10.714% |   15.476% |   47.619% |   15.476% |   10.714% |    6.022% | 
|                 |    4.186% |    6.771% |    8.333% |    3.210% |    8.738% |           | 
|                 |    0.645% |    0.932% |    2.867% |    0.932% |    0.645% |           | 
|-----------------|-----------|-----------|-----------|-----------|-----------|-----------|
|               3 |       14  |       26  |       83  |       55  |       14  |      192  | 
|                 |   29.591  |   26.426  |   66.065  |   55.742  |   14.176  |           | 
|                 |    8.215  |    0.007  |    4.341  |    0.010  |    0.002  |           | 
|                 |    7.292% |   13.542% |   43.229% |   28.646% |    7.292% |   13.763% | 
|                 |    6.512% |   13.542% |   17.292% |   13.580% |   13.592% |           | 
|                 |    1.004% |    1.864% |    5.950% |    3.943% |    1.004% |           | 
|-----------------|-----------|-----------|-----------|-----------|-----------|-----------|
|               4 |       17  |       25  |      108  |       97  |       20  |      267  | 
|                 |   41.151  |   36.748  |   91.871  |   77.516  |   19.714  |           | 
|                 |   14.174  |    3.756  |    2.832  |    4.897  |    0.004  |           | 
|                 |    6.367% |    9.363% |   40.449% |   36.330% |    7.491% |   19.140% | 
|                 |    7.907% |   13.021% |   22.500% |   23.951% |   19.417% |           | 
|                 |    1.219% |    1.792% |    7.742% |    6.953% |    1.434% |           | 
|-----------------|-----------|-----------|-----------|-----------|-----------|-----------|
|               5 |       17  |       26  |      112  |      198  |       19  |      372  | 
|                 |   57.333  |   51.200  |  128.000  |  108.000  |   27.467  |           | 
|                 |   28.374  |   12.403  |    2.000  |   75.000  |    2.610  |           | 
|                 |    4.570% |    6.989% |   30.108% |   53.226% |    5.108% |   26.667% | 
|                 |    7.907% |   13.542% |   23.333% |   48.889% |   18.447% |           | 
|                 |    1.219% |    1.864% |    8.029% |   14.194% |    1.362% |           | 
|-----------------|-----------|-----------|-----------|-----------|-----------|-----------|
|    Column Total |      215  |      192  |      480  |      405  |      103  |     1395  | 
|                 |   15.412% |   13.763% |   34.409% |   29.032% |    7.384% |           | 

\begin{center}

Total Observations in Table: 1395

\end{center}

Figure 3. Cross tabulation table. See appendix for key.

\newpage

|Pearson's Chi-squared test|            |                   |
|--------------------------|------------|-------------------|
|Chi^2 =  482.9719         | d.f. =  20 | p =  1.063345e-89 |
Figure 4. Chart. Pearson's chi-squared test.

### Section 7: Appendix

\begin{tabular}{|l|}
\hline
\textbf{Cell Contents Key for cross tabulation table} \\ \hline
Count                              \\ \hline
Expected Values                    \\ \hline
Chi-square contribution            \\ \hline
Row Percent                        \\ \hline
Column Percent                     \\ \hline
Total Percent                      \\ \hline
\end{tabular}
