---
title: "Unit 10 Lab Report"
author: "Christopher Byrd"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: lualatex
    highlight: pygments
    toc: true
    toc_depth: 3
mainfont: Times New Roman
---

```{r setupChunk, warning=FALSE, include=FALSE}
library(tidyverse)
library(extrafont)
library(ggplot2)
library(ggthemes)
library(hrbrthemes)
library(thematic)
library(extrafontdb)
library(readxl)
library(colorspace)
library(addinslist)
library(clipr)
library(gmodels)
library(Hmisc)
library(RColorBrewer)
library(DescTools)
library(viridis)

Module_10_Lab_Data <- read_excel("data/Module 10 Lab Data.xlsx", 
                                 col_types = c("numeric", "numeric"))

rsq <- function (x, y) cor(as.matrix(x), as.matrix(y)) ^ 2
```

### Section 1: Hypothesis

The goal of this research paper is to determine if age of those who pray 
regularly has a statistically significant effect on income. The 
following are the hypothesizes examined:

> Null: There is no significant correlation between age of those who
> pray regularly and their income.\
> Research~1~: There is a significant correlation between age of those 
> who pray regularly and their income.\

> H~0~: ρ~xy~ = 0\
> H~1~: r~xy~ ≠ 0

\newpage
### Section 2: Data

The data being utilized is from the Baylor Survey of Religious Life.
Participants who prayed at least once or more a day were asked to
respond with their age and the bracket in which their income falls 
within. The following code book was used:

| Question or Description    |        Possible Answers           |
|----------------------------|-----------------------------------|
| Age range of people who    | Range: 17-95                      |
| pray once or several times |                                   |
| a day as reported in 2017. |                                   |
|--------------------------- |-----------------------------------|
| By your best estimate,     | 1) $10,000 or less                |
| what was your total        | 2) $10,001 to $20,000             |
| household income last year,| 3) $20,001 to $35,000             |
| before taxes?              | 4) $35,001 to $50,000             |
|                            | 5) $50,001 to $100,000            |
|                            | 6) $100,001 to $150,000           |
|                            | 7) $150,001 or more               |

### Section 3: Method

To evaluate the data, Pearson's r was calculated using `rcorr`, a
function which evaluates Pearson's r and the associated P-value. The 
data needed to be coerced into a matrix to allow calculation of each
variable. The other option that exists is to create two separate data
frames for each variable.

```{r warning=FALSE, results='hide', eval=FALSE}
rcorr(as.matrix(Module_10_Lab_Data), type = "pearson")
```

To determine \(R^2\), I took a function off the internet to create a 
function for evaluating \(R^2\). The original function required you to 
coerce the variables to be matrices, so I put it in the function to 
avoid this.

```{r warning=FALSE, results='hide', eval=FALSE}
rsq <- function (x, y) cor(as.matrix(x), as.matrix(y)) ^ 2

rsq(Module_10_Lab_Data, Module_10_Lab_Data)
```

Lastly, ggplot2 was used to produce a scatter plot. This was achieved
through the geom `geom_point()`, while `geom_smooth()` was used to create
the trend line.

```{r warning=FALSE, results='hide', eval=FALSE}
graphdata <- as.data.frame(Module_10_Lab_Data)

ggplot(data = graphdata, aes(x = Age,y = `Income Interval`)) +
  geom_point(shape = 18, color = "black") +
  geom_smooth(method = lm, color = "#f21f1f") +
  xlim(0, 100) +
  ylim(0, 8) +
  theme_bw()
```

### Section 4: Results

Due to there being only 2 variables, each individual result within each
statistic is identical. This is because evaluating Age-Income is
identical to Income-Age for r and \(R^2\).

As for the results, Pearson's r is \(-0.11\), showing a weak to 
nonexistent negative correlation between both variables. The p-value for 
this test was \(0.0055\), showing this result to likely be statistically
insignificant.

Lastly, for \(R^2\), the result was 0.0131723. This shows a the that the
model visible from the trend line in figure 1 is a poor predictor and
does not explain and predict the variability observed.

### Section 5: Discussion and Conclusion

From these results, it's clear that age has little to no impact on
income.

It stands to reason the null is accepted and H~1~ is rejected.

### Section 6: Graphics and Visuals

```{r graph, warning=FALSE, echo=FALSE, fig.cap="Scatter plot with trend line. Age and income of respondents."}
graphdata <- as.data.frame(Module_10_Lab_Data)

ggplot(data = graphdata, aes(x = Age,y = `Income Interval`)) +
  geom_point(shape = 18, color = "black") +
  geom_smooth(method = lm, color = "#f21f1f") +
  xlim(0, 100) +
  ylim(0, 8) +
  theme_bw()
```

```{r rcorr, warning=FALSE, echo=FALSE}
rcorr(as.matrix(Module_10_Lab_Data), type = "pearson")
```

\begin{center}
Figure 2: Matrix chart with n value. Pearson's r and p-values of age and income interval.
\end{center}

```{r rsq, echo=FALSE, warning=FALSE}
rsq(Module_10_Lab_Data, Module_10_Lab_Data)
```

\begin{center}
Figure 3: Matrix chart. \(r^2\) of age and income interval.
\end{center}