---
title: "Unit 10 Lab Report"
author: "Christopher Byrd"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: lualatex
    highlight: pygments
    toc: true
    toc_depth: 3
mainfont: Times New Roman
---

```{r setupChunk, warning=FALSE, include=FALSE}
library(tidyverse)
library(extrafont)
library(ggplot2)
library(ggthemes)
library(hrbrthemes)
library(thematic)
library(extrafontdb)
library(readxl)
library(colorspace)
library(addinslist)
library(clipr)
library(gmodels)
library(Hmisc)
library(RColorBrewer)
library(DescTools)
library(viridis)

Module_10_Lab_Data <- read_excel("data/Module 10 Lab Data.xlsx", 
                                 col_types = c("numeric", "numeric"))
```

### Section 1: Hypothesis

The goal of this research paper is to determine if regular prayer has
any correlation on the opinion of the federal government allowing prayer 
in school. The following are the hypothesizes examined:

> Null: There is no significant correlation between prayer frequency and
> opinion on prayer in school.\
> Research~1~: There will be a significance between prayer frequency and
> opinion on prayer in schools.

> H~0~: ρ~xy~ = 0\
> H~1~: r~xy~ ≠ 0

\newpage
### Section 2: Data

The data being utilized is from the Baylor Survey of Religious Life.
Participants were asked to respond with their age and the bracket to which
their income falls within. The following code book was used:

| Question or Description    |        Possible Answers           |
|----------------------------|-----------------------------------|
| Age range of people who    | Range: 17-95                      |
| pray once or several times |                                   |
| a day as reported in 2017. |                                   |
|--------------------------- |-----------------------------------|
| By your best estimate,     | 1) $10,000 or less                |
| what was your total        | 2) $10,001 to $20,000             |
| household income last year,| 3) $20,001 to $35,000             |
| before taxes?              | 4) $35,001 to $50,000             |
|                            | 5) $50,001 to $100,000            |
|                            | 6) $100,001 to $150,000           |
|                            | 7) $150,001 or more               |

### Section 3: Method

To evaluate the data, Pearson's r was calculated using `rcorr`, a
function which evaluates Pearson's r and and the associated P-value. The
data needed to be coerced into a matrix to allow calculation of each
variable. The other option that exists is to create two separate data
frames for each variable.

```{r warning=FALSE, results='hide', eval=FALSE}
rcorr(as.matrix(Module_10_Lab_Data), type = "pearson")
```

To determine R^2^, I took a function off the internet to create a 
function for evaluating R^2^. The original function required you to 
coerce the variables to be matrices, so I put it in the function to 
avoid this.

```{r warning=FALSE, results='hide', eval=FALSE}
rsq <- function (x, y) cor(as.matrix(x), as.matrix(y)) ^ 2

rsq(Module_10_Lab_Data, Module_10_Lab_Data)
```

Lastly, ggplot2 was used to produce a scatter plot. This was achieved
through the geom `geom_point()`, while `geom_smooth()` was used to create
the trend line.

```{r warning=FALSE, results='hide', eval=FALSE}
graphdata <- as.data.frame(Module_10_Lab_Data)

ggplot(data = graphdata, aes(x = Age,y = `Income Interval`)) +
  geom_point(shape = 18, color = "black") +
  geom_smooth(method = lm, color = "#f21f1f") +
  xlim(0, 100) +
  ylim(0, 8) +
  theme_bw()
```

### Section 4: Results

The results from `CrossTable()` can be viewed in full in figure 3. As for
the results of the Chi-squared test (fig. 4), it returns a p-value of 
1.06E-89, well into statistical significance. This is further
demonstrated in figure 1.

### Section 5: Discussion and Conclusion

From these results, it's clear that prayer frequency has a direct and 
measurable impact on opinion regarding the allowance of prayer by the
federal government in schools.

It stands to reason the null is rejected and H~1~ is accepted.

### Section 6: Graphics and Visuals

```{r warning=FALSE, echo=FALSE, fig.cap="Scatter plot with trend line. Age and income of respondents."}
graphdata <- as.data.frame(Module_10_Lab_Data)

ggplot(data = graphdata, aes(x = Age,y = `Income Interval`)) +
  geom_point(shape = 18, color = "black") +
  geom_smooth(method = lm, color = "#f21f1f") +
  xlim(0, 100) +
  ylim(0, 8) +
  theme_bw()
```

```{r warning=FALSE, echo=FALSE}
rcorr(as.matrix(Module_10_Lab_Data), type = "pearson")
```
