---
title: "Unit 10 Lab Report"
author: "Christopher Byrd"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: lualatex
    highlight: pygments
    toc: true
    toc_depth: 3
mainfont: Times New Roman
---

```{r setupChunk, warning=FALSE, include=FALSE}
library(tidyverse)
library(extrafont)
library(ggplot2)
library(ggthemes)
library(hrbrthemes)
library(thematic)
library(extrafontdb)
library(readxl)
library(colorspace)
library(addinslist)
library(clipr)
library(gmodels)
library(Hmisc)
library(RColorBrewer)
library(DescTools)
library(viridis)

Module_10_Lab_Data <- read_excel("data/Module 10 Lab Data.xlsx", 
                                 col_types = c("numeric", "numeric"))
```

### Section 1: Hypothesis

The goal of this research paper is to determine if regular prayer has
any correlation on the opinion of the federal government allowing prayer 
in school. The following are the hypothesizes examined:

> Null: There is no significant correlation between prayer frequency and
> opinion on prayer in school.\
> Research~1~: There will be a significance between prayer frequency and
> opinion on prayer in schools.

> H~0~: ρ~xy~ = 0
> H~1~: r~xy~ ≠ 0

### Section 2: Data

### Section 3: Method

```{r warning=FALSE, results='hide', eval=FALSE}
CrossTable(Module_9_Lab_Data$PrayerFreq, Module_9_Lab_Data$PrayerSchool,
           expected = TRUE, dnn = c("Prayer Frequency",
                                    "Prayer in School"),
           format = c("SPSS"))
```

### Section 4: Results

The results from `CrossTable()` can be viewed in full in figure 3. As for
the results of the Chi-squared test (fig. 4), it returns a p-value of 
1.06E-89, well into statistical significance. This is further
demonstrated in figure 1.

### Section 5: Discussion and Conclusion

From these results, it's clear that prayer frequency has a direct and 
measurable impact on opinion regarding the allowance of prayer by the
federal government in schools.

It stands to reason the null is rejected and H~1~ is accepted.

### Section 6: Graphics and Visuals