---
title: "Lab 8 Report"
author: "Christopher Byrd"
date: "2024-11-11"
output:
  html_document: default
  pdf_document: default
---

```{r setupChunk, warning=FALSE, include=FALSE}
library(plyr)
library(DescTools)
library(ggplot2)
library(dplyr)
library(RColorBrewer)
library(hrbrthemes)
library(tidyverse)
library(viridis)
library(forcats)
library(readxl)
library(knitr)
library(addinslist)
library(clipr)


Module_8_Lab_Data <- read_excel(
  "C:/Users/byrds/Downloads/Module 8 Lab Data.xlsx")
```

### Section 1: Hypothesis

The goal of this research paper is to determine if regular prayer has
any correlation to religious service attendance. Also examined is the
affect sex and prayer has in relation to service attendance. The
following are the hypothesizes examined:

> Null: There is no significant correlation between prayer frequency,
> gender, and attendance of religious services.\
> Research~1~: There will be a significant increase in attendance with
> an increase in prayer frequency\
> Research~2~: Women will attend religious services more with higher
> levels of prayer when compared to men.

> H~0~: Higher prayer frequency + gender =/= More religious attendance\
> H~1~: Higher prayer frequency = More religious attendance\
> H~2~: Women = More religious attendance

### Section 2: Data

The data used for this report is from the Baylor Survey of Religious
Life. The data records three different metrics of each respondent:
frequency of prayer, gender, and attendance of religious services. Data
is encoded as ordinal for frequency and attendance, while gender is
nominal.

### Section 3: Method

Using RStudio, data was first packaged within a `data.frame` to make for
easier coding, as the `data.frame` it came prepackaged in had an
excessively long name.

```{r warning=FALSE, results='hide'}
betterList <- data.frame(prayerFrequency = Module_8_Lab_Data$PrayerFreq,
                         attendance = Module_8_Lab_Data$Attendance,
                         gender = Module_8_Lab_Data$Gender)
```

#### Section 3.1: Prayer Frequency\

An ANOVA (`aov()`) was performed with respects to prayer frequency and 
religious service attendance. A `TukeyHSD()` was performed to determine 
where variance might be resulting from.

```{r warning=FALSE, results='hide'}
# Prayer and Attendance

storedVal <- aov(betterList$attendance~betterList$prayerFrequency,
                 betterList)
summary(storedVal)

TukeyHSD(aov(betterList$attendance~as.factor(betterList$prayerFrequency),
             betterList))
```

#### Section 3.2: Prayer and Gender Frequency\

`lm()` was used to perform an analysis (factorial ANOVA) of the 
interaction between gender and prayer frequency and how both affect 
attendance of religious services. An ANOVA was used to perform a 
`TukeyHSD()` test. Note that results of the `aov()` and `lm()` are 
similar and within margin of error; thus it seems reasonable to use 
`aov()` to construct a `TukeyHSD`, as there is no applicable method 
within `TukeyHSD` to run the test with `lm()` as the input. Furthermore,
documentation within the *stats* package reveals that `aov()` is a 
wrapper for `lm()` with differing output methods; therefore, results
should be near identical no matter the scenario.

```{r warning=FALSE, results='hide'}
# Gender, Prayer, and Attendance

storedVal <- lm(betterList$attendance~betterList$prayerFrequency*betterList$gender,
                data = betterList)
summary(storedVal)

TukeyHSD(aov(betterList$attendance~as.factor(betterList$prayerFrequency)*as.factor(betterList$gender),
             data = betterList))
```

Lastly, *ggplot2* was used to create a interaction plot between service
attendance in relation to gender and prayer frequency.

```{r warning=FALSE, results='hide', fig.keep='none'}
# ggplot2 `interaction.plot()`

listInt <- ddply(betterList,.(prayerFrequency,gender),summarise, val = mean(attendance))

ggplot() +
  geom_point(data = betterList, aes(x = prayerFrequency, y = attendance, colour = gender)) +
  geom_point(data = listInt, aes(x = prayerFrequency, y = val, colour = gender)) +
  geom_line(data = listInt, aes(x = prayerFrequency, y = val, group = gender, colour = gender)) +
  ylim(1, 5) +
  theme_bw() +
  labs(title = "Interaction Plot", x = "Prayer Frequency", y = "Mean of Attendance")
```

### Section 4: Results

From these tests, a number of discoveries emerge. The data is as 
follows:

#### Section 4.1: Attendance in Relation to Prayer\

1. ANOVA analysis ~
```{r echo=FALSE, results='markup'}
# Gender, Prayer, and Attendance
storedVal <- aov(betterList$attendance~betterList$prayerFrequency,
                 betterList)
summary(storedVal)
```

2. `TukeyHSD()` ~
```{r echo=FALSE, results='markup'}
TukeyHSD(aov(betterList$attendance~as.factor(betterList$prayerFrequency),
             betterList))
```

#### Section 4.2: Attendance in Relation to Sex and Prayer:\

1. `lm()` analysis ~

```{r echo=FALSE, results='markup'}
# Gender, Prayer, and Attendance

storedVal <- lm(betterList$attendance~betterList$prayerFrequency*betterList$gender,
                data = betterList)
summary(storedVal)
```

2. `TukeyHSD()` using `AOV()` ~

```{r echo=FALSE, results='markup'}
TukeyHSD(aov(betterList$attendance~as.factor(betterList$prayerFrequency)*as.factor(betterList$gender),
             data = betterList))
```

3. Interaction plot: **See Section 6, Fig. 1.**

#### Section 4.3: Key Values\

Taking what is most important for prayer frequency, there is a 
statistically significant link between prayer and attendance. With an 
f-statistic of 733.6, there is no possible combinations of degrees of 
freedom in which the results, according to the critical f-value, are 
not significant (Fig. 2). The p-value for this test is 2E-16, or 
0.0000000000000002. *Fig. 1* represents this graphically by way of mean 
of attendance based on prayer rates. The Tukey Post Hoc test 
(`TukeyHSD()`) reveals that the interaction of each level between both 
factors is significant, with p-values listed as 0.\

For attendance in relation to gender and prayer, we see a mix of 
statistical significance in interaction between genders and prayer 
frequency. Overall, the findings within the `lm()` analysis shows
prayer frequency has a statistically significant effect on service 
attendance. However, gender does not have this effect. The individual 
interactions displayed in the Tukey Post Hoc test corroborate this.

### Section 5: Discussion and Conclusion

From these results, it's clear that prayer frequency has a direct and 
measurable impact on attendance of religious services. Gender, however 
has no significant impact, and any seen is likely due to error. This is 
further confirmed by *fig. 1*, in which we can see that the attendance 
results for each gender according to prayer frequency show similar 
trends.

It stands to reason the null is rejected and H~1~ is accepted. However,
H~2~ is rejected on the grounds of not possessing statistical 
significance.

### Section 6: Graphics and Visuals

```{r echo=FALSE, warning=FALSE, fig.cap="Fig. 1. Line plot. Effect of gender and prayer frequency on mean of attendance."}
# ggplot2 `interaction.plot()`

listInt <- ddply(betterList,.(prayerFrequency,gender),summarise, val = mean(attendance))

ggplot() +
  geom_point(data = betterList, aes(x = prayerFrequency, y = attendance, colour = gender)) +
  geom_point(data = listInt, aes(x = prayerFrequency, y = val, colour = gender)) +
  geom_line(data = listInt, aes(x = prayerFrequency, y = val, group = gender, colour = gender)) +
  ylim(1, 5) +
  theme_bw() +
  labs(title = "Interaction Plot", x = "Prayer Frequency", y = "Mean of Attendance", 
       colour = "Gender")
```

|                         | Df     | Sum Sq |Mean Sq| F value | Pr(>F)     |
|-------------------------|--------|--------|-------|---------|------------|
| Prayer Frequency        | 1      | 3166   | 3166  | 733.2   | <2e-16 *** |
| Residuals               | 1398   | 6037   | 4     | 4       |            |
Fig. 2. F table for Prayer Frequency.\

| F-statistic: | Df~denom~:| Df~num~:  | p-value:  |
|--------------|-----------|-----------|-----------|
| 244.6        | 3         | 1396      | < 2.2e-16 |

Fig. 3. Linear Model table for Prayer and Gender.