---
title: "Unit 11 Lab Report"
author: "Christopher Byrd"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: lualatex
    highlight: pygments
    toc: true
    toc_depth: 3
mainfont: Times New Roman
---

```{r setupChunk, warning=FALSE, include=FALSE}
library(tidyverse)
library(extrafont)
library(ggplot2)
library(ggthemes)
library(hrbrthemes)
library(thematic)
library(extrafontdb)
library(readxl)
library(colorspace)
library(addinslist)
library(clipr)
library(gmodels)
library(Hmisc)
library(RColorBrewer)
library(DescTools)
library(viridis)

Module_11_Lab_Data <- read_excel("data/Module 11 Lab Data.xlsx", 
                                 col_types = c("numeric", "numeric",
                                               "numeric", "numeric"))
```

### Section 1: Hypothesis

The goal of this research paper is to determine how age, income, and 
regular attendance of church services affects prayer rates.

> Null: There is no significant correlation between age of those who
> pray regularly and their income, age, and church attendance.\
> Research~1~: There is a significant correlation between age of those 
> who pray regularly and their income, age, and church attendance.

> H~0~: ρ~xy~ = 0\
> H~1~: r~xy~ ≠ 0

\newpage
### Section 2: Data

The data being utilized is from the Baylor Survey of Religious Life.
Participants asked to respond to the question listed in the following 
code book:

| Question or Description    |        Possible Answers           |
|----------------------------|-----------------------------------|
| Age range of people who    | Range: 17-95                      |
| pray once or several times |                                   |
| a day as reported in 2017. |                                   |
|--------------------------- |-----------------------------------|
| By your best estimate,     | 1) $10,000 or less                |
| what was your total        | 2) $10,001 to $20,000             |
| household income last year,| 3) $20,001 to $35,000             |
| before taxes?              | 4) $35,001 to $50,000             |
|                            | 5) $50,001 to $100,000            |
|                            | 6) $100,001 to $150,000           |
|                            | 7) $150,001 or more               |
|----------------------------|-----------------------------------|
| About how often do you     | 0) Never                          |
| spend time alone praying   | 1) Only on certain occasions      |
| outside of religious       | 2) Once a week or less            |
| services?                  | 3) A few times a week             |
|                            | 4) Once a day                     |
|                            | 5) Several times a day            |
|----------------------------|-----------------------------------|
| Attends a worship service  | 1) No                             |
| about once a week or more. | 2) Yes                            |

Attendance was originally a nominal variable. To allow for a linear 
regression, it was recoded as a dummy variable.

### Section 3: Method

To evaluate the data, the data set was first re-coded with shorter 
variable names, mainly because of my hate for long variable names. The
new data set was then coerced to be a `data.frame` to make later
actions easier (particularly with `ggplot2`, since data needs to be in 
tidy format).

```{r warning=FALSE, results='hide', eval=TRUE}
# Re-coded with shorter names because idk I don't like long variable
# names. ¯\_(ツ)_/¯

pf <- Module_11_Lab_Data$`Pray Frequency`
inc <- Module_11_Lab_Data$`Income Intervals`
age <- Module_11_Lab_Data$AGE
regat <- Module_11_Lab_Data$`Regular Attendance`

df <- data.frame(pf, inc, age, regat)
```

The next evaluation consisted of finding the median of income interval,
age, and regular attendance. To cut down on the amount of code I needed
to write, I made this a `for` loop. From here, the linear model (AKA 
regression) was calculated with prayer frequency regressed over income
interval, age, and regular attendance of church services.

```{r warning=FALSE, results='hide', eval=FALSE}
for (i in 2:4) {
  print(paste(median(df[[i]])))
}

lmresult <- lm(formula = pf ~ inc + age + regpr, df)
```

After running a summary to find the exact results of the `lm()`, a
ggplot2 plot was created based upon the fitted values.

I did all this work below thinking I was working with Pearson's r, but
I wasn't. I just put it here to display the work I put into this.

```{r eval=FALSE, warning=FALSE, results='hide'}
for (i in 1:length(lmresult$coefficients)) {
  if (lmresult$coefficients[i] > 0) {
    print(paste("The coefficient for", names(lmresult$coefficients[i]),
                "is", lmresult$coefficients[i], "and direct"))
  }   else {
    print(paste("The coefficient for", names(lmresult$coefficients[i]),
                "is", lmresult$coefficients[i], "and indirect"))
  }
  if (abs(lmresult$coefficients[i]) > 0.2) {
    print("There is a weak to strong relationship.")
  }   else {
    print("There is little to no relationship")
  }
}
```

### Section 4: Results



### Section 5: Discussion and Conclusion



### Section 6: Graphics and Visuals

```{r graph, warning=FALSE, echo=FALSE, results='hide', fig.cap="Scatter plot with linear regression based upon fitted values."}
```

\begin{center}
Figure 3: Matrix chart. \(r^2\) of age and income interval.
\end{center}