---
title: "Unit 11 Lab Report"
author: "Christopher Byrd"
date: "`r Sys.Date()`"
output:
  pdf_document:
    latex_engine: lualatex
    highlight: pygments
    toc: true
    toc_depth: 3
mainfont: Times New Roman
---

```{r setupChunk, warning=FALSE, include=FALSE}
library(tidyverse)
library(extrafont)
library(ggplot2)
library(ggthemes)
library(hrbrthemes)
library(thematic)
library(extrafontdb)
library(readxl)
library(colorspace)
library(addinslist)
library(clipr)
library(gmodels)
library(Hmisc)
library(RColorBrewer)
library(DescTools)
library(viridis)
library(ggpmisc)

Module_11_Lab_Data <- read_excel("data/Module 11 Lab Data.xlsx", 
                                 col_types = c("numeric", "numeric",
                                               "numeric", "numeric"))
```

### Section 1: Hypothesis

The goal of this research paper is to determine how age, income, and 
regular attendance of church services affects prayer rates.

> Null: There will be no significant difference in prayer frequency
> based upon age, income range, or church attendance\
> Research~1~: There will be a significant difference in prayer 
> frequency based upon age, income range, or church attendance.\
> Research~2~: Median age, income range, and church attendance will 
> result in a prayer rate of a few times a week.

> H~0~: ρ~xy~ = 0\
> H~1~: r~xy~ ≠ 0\
> H~2~: Y~x_median~ = 3

\newpage
### Section 2: Data

The data being utilized is from the Baylor Survey of Religious Life.
Participants asked to respond to the question listed in the following 
code book:

| Question or Description    |        Possible Answers           |
|----------------------------|-----------------------------------|
| Age range of people who    | Range: 17-95                      |
| pray once or several times |                                   |
| a day as reported in 2017. |                                   |
|--------------------------- |-----------------------------------|
| By your best estimate,     | 1) $10,000 or less                |
| what was your total        | 2) $10,001 to $20,000             |
| household income last year,| 3) $20,001 to $35,000             |
| before taxes?              | 4) $35,001 to $50,000             |
|                            | 5) $50,001 to $100,000            |
|                            | 6) $100,001 to $150,000           |
|                            | 7) $150,001 or more               |
|----------------------------|-----------------------------------|
| About how often do you     | 0) Never                          |
| spend time alone praying   | 1) Only on certain occasions      |
| outside of religious       | 2) Once a week or less            |
| services?                  | 3) A few times a week             |
|                            | 4) Once a day                     |
|                            | 5) Several times a day            |
|----------------------------|-----------------------------------|
| Attends a worship service  | 1) No                             |
| about once a week or more. | 2) Yes                            |

Attendance was originally a nominal variable. To allow for a linear 
regression, it was recoded as a dummy variable.

### Section 3: Method

To evaluate the data, the data set was first re-coded with shorter 
variable names, mainly because of my hate for long variable names. The
new data set was then coerced to be a `data.frame` to make later
actions easier (particularly with `ggplot2`, since data needs to be in 
tidy format).

```{r warning=FALSE, results='hide', eval=TRUE}
# Re-coded with shorter names because idk I don't like long variable
# names. ¯\_(ツ)_/¯

pf <- Module_11_Lab_Data$`Pray Frequency`
inc <- Module_11_Lab_Data$`Income Intervals`
age <- Module_11_Lab_Data$AGE
regat <- Module_11_Lab_Data$`Regular Attendance`

df <- data.frame(pf, inc, age, regat)
```

The next evaluation consisted of finding the median of income interval,
age, and regular attendance. To cut down on the amount of code I needed
to write, I made this a `for` loop. From here, the linear model (AKA 
regression) was calculated with prayer frequency regressed over income
interval, age, and regular attendance of church services. A multiple 
regression was utilized to show how each variable interacts with prayer 
frequency together.

```{r warning=FALSE, results='hide', eval=FALSE}
for (i in 2:4) {
  print(paste(median(df[[i]])))
}

lmresult <- lm(formula = pf ~ inc + age + regpr, df)
```

After running a summary to find the exact results of the `lm()`, a
ggplot2 plot was created based upon the fitted values.

```{r warning=FALSE, results='hide', eval=FALSE}
df$fitted <- lmresult$fitted.values
df$residuals <- lmresult$residuals

ggplot(df, aes(x = fitted, y = pf)) +
  geom_point() +
  geom_smooth(method = 'lm', color = '#679df5') +
  labs(title = "Prayer Frequency Based Upon Fitted Values") +
  xlab("Fitted values (see section 2-4 for more info)") +
  ylab("Prayer Frequency") +
  stat_poly_eq(use_label(c("eq", "R2")), rr.digits = 4,vjust = 1.3) +
  theme_pander() +
  theme(plot.margin = margin(t = 15, r = 15, b = 15, l = 15, unit = "pt")) +
  geom_point(aes(x = x_value, y = y_value), color = "red", size = 2.2) +
  geom_text(aes(x = x_value, y = y_value,
                label = paste("(", round(x_value, 2), ", ",
                              round(y_value, 2), ")", sep = "")),
            vjust = -1, hjust = 1, size = 3.2)
```

I completed the work below thinking I was working with Pearson's r, which
I wasn't. I've placed it here to display the work I put into it:

```{r eval=FALSE, warning=FALSE, results='hide'}
for (i in 1:length(lmresult$coefficients)) {
  if (lmresult$coefficients[i] > 0) {
    print(paste("The coefficient for", names(lmresult$coefficients[i]),
                "is", lmresult$coefficients[i], "and direct"))
  }   else {
    print(paste("The coefficient for", names(lmresult$coefficients[i]),
                "is", lmresult$coefficients[i], "and indirect"))
  }
  if (abs(lmresult$coefficients[i]) > 0.2) {
    print("There is a weak to strong relationship.")
  }   else {
    print("There is little to no relationship")
  }
}
```

### Section 4: Results

Full results can be found in figure 2 within section 6. Figure 2 shows that
each individual independent variable is statistically significant in relation
to how they affect the dependent. Furthermore, regular prayer and age display
positive relationships, while income displays a negative relationship. Regular
prayer is the largest predictor and the most statistically significant, at
a p-value of \(2E-16\).

Medians for the 3 dependent variables are as follows:

```{r median, echo=FALSE, warning=FALSE, results='markup'}
for (i in 2:4) {
  print(paste(median(df[[i]])))
}
```

...while median for the fitted values is *2.241326*, based upon the following string: `median(df$fitted)`. To find the median of prayer frequency, the 
following code snippet was used:

```{r, echo=TRUE, warning=FALSE}
x_value <- 2.241326
y_value <- (2.72e-14) + 1 * x_value
print(y_value)
```

As is shown, the value is 2.24, which, when rounded, means median prayer 
frequency falls within once a week or less. Fitted values are plotted in
figure 1.

### Section 5: Discussion and Conclusion

From these results, the null is rejected and research~1~ is accepted.
Research~2~ is rejected, as the actual value of Y is 2. 
\newpage

### Section 6: Graphics and Visuals

```{r graph, warning=FALSE, echo=FALSE, results='hide', fig.cap="Scatter plot with linear regression based upon fitted values. Median value of x labeled."}
lmresult <- lm(formula = pf ~ inc + age + regat, df)

df$fitted <- lmresult$fitted.values
df$residuals <- lmresult$residuals

x_value <- 2.241326
y_value <- (2.72e-14) + 1 * x_value

ggplot(df, aes(x = fitted, y = pf)) +
  geom_point() +
  geom_smooth(method = 'lm', color = '#679df5') +
  labs(title = "Prayer Frequency Based Upon Fitted Values") +
  xlab("Fitted values (see section 2-4 for more info)") +
  ylab("Prayer Frequency") +
  stat_poly_eq(use_label(c("eq", "R2")), rr.digits = 4,vjust = 1.3) +
  theme_pander() +
  theme(plot.margin = margin(t = 15, r = 15, b = 15, l = 15, unit = "pt")) +
  geom_point(aes(x = x_value, y = y_value), color = "red", size = 2.2) +
  geom_text(aes(x = x_value, y = y_value,
                label = paste("(", round(x_value, 2), ", ",
                              round(y_value, 2), ")", sep = "")),
            vjust = -1, hjust = 1, size = 3.2)
```

```{r Regression Table, echo=FALSE, warning=FALSE, results='hold'}
summary(lmresult)
```

\begin{center}
Figure 2: Regression Table. Test statistics included.
\end{center}